---
title: "PA1_template.Rmd"
author: "ManSab"
date: "6/17/2019"
output: html_document
---
### Activity Monitoring Peer-graded Assignment
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, results = "markup", dev="png" )
library("lubridate")
library("ggplot2")
library("magrittr")
library("dplyr")
```

####1) Read activity file into data frame that has following columns: steps, date and interval. Date is also formated.
``` {r}

Activity <- read.csv("activity.csv")
Activity$date <- as.character(Activity$date)
Activity$date <- ymd(Activity$date)
```
####2) Histogram of the total number of steps taken each day
``` {r dpi=36, out.width="400px", out.height="400px"}
Act_Hist <- Activity %>% group_by(date) %>% summarise(st = sum(steps))
#graphics.off()
hist(Act_Hist$st,xlab = "Daily Steps", main = "Daily Steps", col = "RED")
```

####3) Mean and Median steps taken each day
```{r}
summary(Act_Hist)
```

####4) Time series plot of the average number of steps taken
Make a time series plot (i.e.type="l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r dpi=36, out.width="400px", out.height="400px"}
Act_Avg <- aggregate(Activity$steps~Activity$interval,Activity,mean)
plot(Act_Avg$`Activity$steps`~Act_Avg$`Activity$interval`, type="l", xlab = "Time Interval", ylab = "Avg.Steps",main = "Average steps", col = "blue")
```

####5) The 5-minute interval that, on average, contains the maximum number of steps
```{r}

Act_Avg[which.max(Act_Avg$`Activity$steps`),]
```
####6) Code to describe and show a strategy for imputing missing data.
Let's check NA's in the data set

```{r}
colSums(is.na(Activity))
```
There are 2304 missing(NA) values in steps column of the data set Activity. Let's replace column 'Step' missing(NA) values with mean for time-interval average.
```{r}
ImpActivity <- Activity
for(i in 1:nrow(ImpActivity)){
    if (is.na(ImpActivity$steps[i])){
        inter <- ImpActivity$interval[i]
        capstep <- Act_Avg[Act_Avg$`Activity$interval`==inter,]
        ImpActivity$steps[i] <- capstep$`Activity$steps`
    }
}

```
####7) Histogram of the total number of steps taken each day after missing values are imputed.
```{r dpi=36, out.width="400px", out.height="400px"}
ImpHist <- ImpActivity %>% group_by(date) %>% summarise(st = sum(steps))
#graphics.off()
hist(ImpHist$st,xlab = "Daily Steps", main = "Daily Steps", col = "blue")

```

####8) Mean and Median steps taken each day after values are imputed. There clearly are differences in median, first and third quantile values.
```{r}
summary(ImpHist)
```

####9) Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r dpi=36, out.width="400px", out.height="400px"}

ImpHistWD <- ImpActivity %>% mutate(Wtype = if_else(weekdays(as.Date(ImpActivity$date)) %in% c('Saturday','Sunday'),"Weekend","Weekday"))
ImpHistWD$Wtype <- as.factor(ImpHistWD$Wtype)

Dail_Avg <- aggregate(ImpHistWD$steps~ImpHistWD$interval+ImpHistWD$Wtype,ImpHistWD,mean)
#graphics.off()
ggplot(Dail_Avg,aes(x=Dail_Avg$`ImpHistWD$interval`,y=Dail_Avg$`ImpHistWD$steps`),color=`Weekday or Weekend`) +  geom_line() + facet_wrap(.~Dail_Avg$`ImpHistWD$Wtype`, nrow = 2) + xlab("Interval") + ylab("Average number of steps") 

```

*############   Session info for package and other details #############*
```{r}
sessionInfo()
```
